---
#title: "Hypercube Walks"
author: "Sergei Tarasov"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
params:
  output_dir: "/output"
---





```{r, setup,  include=FALSE}
library(DT)
library(ggplot2)
library(knitr)
library(kableExtra)
library(here)
knitr::opts_knit$set(root.dir = here::here(), cache = TRUE)
```

```{r message=FALSE}

source('R/dependencies.R')
source('R/utils.R')

```

# Construct Hypercube

```{r}
Q <- initQ(c(0,1), c(1), diag.as = 0)
Q
Qj <-  amaSMM(Q, Q, Q, Q, Q, Q, Q, Q, Q, Q, controlling.state = NULL)
nrow(Qj)
```

# Plot as grpah

```{r}
# Create graph object
G <- graph.adjacency(Qj, mode = "undirected")
# plot(G, 
#      vertex.size = 5, 
#      edge.width = 2,
#      vertex.label = NA,
#      edge.color = "gray")
```


# Remove states randomly (does not produce rate variation)

```{r}
# we leave the first state, and sample states to remove
# s2remove=sample_states(total=1024, 400)
# Qr <- Qj[-s2remove,-s2remove]
# nrow(Qr)
```

# Select states NON-randomly (this does produce asymmetric variation)

```{r}
#--- mutate
initial_string='0000000000'
selected_states <- c()
# number of replications
for (i in 1:1){
  # number of generations per replication
  x <- mutate_binary_string(initial_string, generations=100)
  selected_states <- c(x, selected_states)
}
selected_states <- unique(selected_states)
length(selected_states)
idx <- match(selected_states, rownames(Qj))
Qr <- Qj[idx,idx]
nrow(Qr)
```

# Mutate strings as unique

```{r}
# initial_string <- '0000000000'
# selected_states <- mutate_unique_binary_string(initial_string, generations=100)
# selected_states <- unique(selected_states)
# length(selected_states)
# idx <- match(selected_states, rownames(Qj))
# Qr <- Qj[idx,idx]
# nrow(Qr)
```


# Make graph and get numbe of connected components

```{r}
Gr <- graph.adjacency(Qr, mode = "undirected")
# Calculate connected components
comp <- components(Gr)
comp$no
```


# Plot graph (=state connectivity) 

```{r}

plot(Gr, 
     vertex.size = 5, 
     edge.width = 2,
     vertex.label = NA,
     edge.color = "gray")

```
# Make rate matrix that is scaled to have wating time for each state = 1

```{r}
Qscl <- adj2Q(Qr) 
head(diag(Qscl))
```

# Calculate approximate rates for each binary character (based on lumpability property)

```{r}
get_lumpQrates_all(Qscl, nchar=10)
```

# Simulate histories

## Simulate tree

```{r}
tree<-pbtree(n=100, scale=100, b=1, d=0)
#plot(tree)

#saveRDS(tree, 'data/tree.rds')
```

## Create Q by scaling rate

```{r}
Qsim=Qscl*.1
diag(Qsim) %>% head

```



## Make history

```{r}
hist <- sim.history(tree, Qsim, nsim=1, anc='0000000000')
plot(hist)

#saveRDS(hist, 'data/hist.rds')
```

## Convert the complex char to binaries

```{r}
#hist$tip.label
#hist$states
tb <- convert_to_binary_chars(hist$states)
tb

#saveRDS(tb, 'data/tb.rds')
```
## Make null history

```{r}
Qscl0 <- adj2Q(Qj) 
#head(diag(Qscl0 ))
Qsim0 <- Qscl0 *0.1
head(diag(Qsim0))
hist0 <- sim.history(tree, Qsim0, nsim=1, anc='0000000000')
plot(hist0)
tb0 <- convert_to_binary_chars(hist0$states)
```

# ML inference

## Read data

```{r}
tree <- readRDS('data/tree.rds')
tb <- readRDS('data/tb.rds')
```


## make Qs

```{r}

Qsym <- initQ(c(0,1), c(1), diag.as = NA)
Qas <- initQ(c(0,1), c(1,2), diag.as = NA)

```


## Inference
```{r}
trait <- cbind(rownames(tb), tb[,3])
res.sym <- rayDISC(tree, trait, rate.mat=Qsym, node.states="none",  model="ARD", root.p="maddfitz")
res.as <- rayDISC(tree, trait, rate.mat=Qas, node.states="none",  model="ARD", root.p="maddfitz")
```

## AIC

```{r}
res.sym$AIC
res.as$AIC
```




